Insert your title here

### 2023-07-19 manuscript_JQAS/manuscript_JQAS.Rnw
JQAS format.

\author*[1]{First Author}
%\ use * to mark the author as the corresponding author
\author[2]{Second Author}
\author[2]{Third Author} 
\runningauthor{F.~Author et al.}
\affil[1]{\protect\raggedright 
Institution, Department, City, Country of first author, e-mail: author\_one@xx.yz}
\affil[2]{\protect\raggedright 
Institution, Department, City, Country of second author and third author, e-mail: author\_two@xx.yz, author\_three@xx.yz}

@@

The first professional baseball team in the United States, the Cinicnnati Red Stockings, was formed in 1869 (@BBHOF1869).   Many leagues came and went in the late 1800s, but National League (NL), formed in 1876, emerged as the predominant league of the time.  A few decades later, the American League (AL) began growing in popularity and eventually reached an agreement with the NL to be the two major leagues of baseball with the winner of each league playing in the World Series starting in 1903.

@@

More specifically, @Woltring2018 mentions six eras of modern baseball": 
"Baseball has endured much change over the course of its history, and because of constant change, the modern era of baseball has been segmented into six distinct sub-eras. A common list presented at Baseball-Reference described the eras as the Dead Ball Era (1901-1919), the Live Ball Era (1920-1941), the Integration Era (1942-1960), the Expansion Era (1961-1976), the Free Agency Era (1977-1993) and the Long Ball/Steroid Era (1994-2005)."  @Woltring2018 notes that they name a seventh era after 2006, which they term the Post Steroid Era.

While many baseball writers have attempted to define the different eras of baseball, there has also been some academic work that has sought to empirically define eras in baseball.  @Groothius2017 looked for structural breaks in univariate times series of performance measures over the period from 1871-2020.  They analyzed four statistics: slugging percentage (SLG, home run (HR) rate, batting average (BA), and runs batted in (RBI) rate.   For each of these statistics, the computed the mean and standard deviation (SD) across all players who had at least 100 at bats in a given season to yield a univariate time series for each of these statistics.  They then used the Lagrange Multiplier (LM) unit root test proposed in @LeeandStrazicich2003 to find structural breaks.  They identified structural breaks in slugging percentage in 1921 and 1992, the first of which marks the end of the Dead Ball Era and the latter corresponding to the start of the steroid era.

@LeeFort2005 looks for structural changes in competitive balance of the two league American and National.  Use methods from Andrews1993, Bai1997, 1999 and Bai and Perron 1998 and 2003 to look for break points between 1901 -1999.  Theuy measure competitive balance in two ways: 1) @Noll1988 and @Scully1989 and 2) Lee 2004. They find break point sin competitive balance in 1912, 1926, and 1933 for the NL and in 1926 and 1957 in the AL.

Baseball is not the only sport where this type of analysis has been applied.  @PalaciosHuerta2004 looked for structural changes in soccer using data from British soccer leagues through 1996.  They had data from division I (Premier League) and II starting in the late 1800s, and data from lower divisions III and IV from just after WWII starting in 1947.  They identify a number of change points.  Notably they identify a change point in the mean of margin of victory in 1925 related to the change in the definition of offsides (changed from 3 players to 2 players), an change points in the variability of number of goals in the early 1980s and 1992 corresponding to the change in number of points for a win (i.e. went from 2 points to 3 points) and a change in the backpass rule, respectively.  @FortLee2007 looked for structural breaks in major North American sports other than baseball (i.e. basketball (NBA), hockey (NHL), and American football (NFL)).  They identified a number of change points related to competitive balance in each sport that often, though not always, correspond to league expansion, league mergers, or other major events in a sport (e.g. increased number of foreign players in the NBA in the late 1990s/early 2000s).

>>>>

\author*[1]{Mena Q.Whalen}
%\ use * to mark the author as the corresponding author
\author[2]{Brian Mills}
\author[3]{Gregory J. Matthews} 
\runningauthor{Whalen et al.}
\affil[1]{\protect\raggedright 
Loyola University Chicago, Department of Mathematics and Statistics, Chicago, IL, USA, e-mail: mwhalen@luc.edu}
\affil[2]{\protect\raggedright 
Texas, Department, City, Country of second author and third author, e-mail: author\_two@xx.yz, author\_three@xx.yz}
\affil[3]{\protect\raggedright 
Loyola University Chicago, Department of Mathematics and Statistics, Chicago, IL, USA, e-mail: gmatthews1@luc.edu}

@@

The first professional baseball team in the United States, the Cinicnnati Red Stockings, was formed in 1869 (\cite{BBHOF1869}).   Many leagues came and went in the late 1800s, but National League (NL), formed in 1876, emerged as the predominant league of the time.  A few decades later, the American League (AL) began growing in popularity and eventually reached an agreement with the NL to be the two major leagues of baseball with the winner of each league playing in the World Series starting in 1903.

@@

More specifically, \cite{Woltring2018} mentions six eras of modern baseball": 
"Baseball has endured much change over the course of its history, and because of constant change, the modern era of baseball has been segmented into six distinct sub-eras. A common list presented at Baseball-Reference described the eras as the Dead Ball Era (1901-1919), the Live Ball Era (1920-1941), the Integration Era (1942-1960), the Expansion Era (1961-1976), the Free Agency Era (1977-1993) and the Long Ball/Steroid Era (1994-2005)."  \cite{Woltring2018} notes that they name a seventh era after 2006, which they term the Post Steroid Era.

While many baseball writers have attempted to define the different eras of baseball, there has also been some academic work that has sought to empirically define eras in baseball.  \cite{Groothius2017} looked for structural breaks in univariate times series of performance measures over the period from 1871-2020.  They analyzed four statistics: slugging percentage (SLG, home run (HR) rate, batting average (BA), and runs batted in (RBI) rate.   For each of these statistics, the computed the mean and standard deviation (SD) across all players who had at least 100 at bats in a given season to yield a univariate time series for each of these statistics.  They then used the Lagrange Multiplier (LM) unit root test proposed in @LeeandStrazicich2003 to find structural breaks.  They identified structural breaks in slugging percentage in 1921 and 1992, the first of which marks the end of the Dead Ball Era and the latter corresponding to the start of the steroid era.

\cite{LeeFort2005} looks for structural changes in competitive balance of the two league American and National.  Use methods from Andrews1993, Bai1997, 1999 and Bai and Perron 1998 and 2003 to look for break points between 1901 -1999.  Theuy measure competitive balance in two ways: 1) \cite{Noll1988} and \cite{Scully1989} and 2) Lee 2004. They find break point sin competitive balance in 1912, 1926, and 1933 for the NL and in 1926 and 1957 in the AL.

Baseball is not the only sport where this type of analysis has been applied.  \cite{PalaciosHuerta2004} looked for structural changes in soccer using data from British soccer leagues through 1996.  They had data from division I (Premier League) and II starting in the late 1800s, and data from lower divisions III and IV from just after WWII starting in 1947.  They identify a number of change points.  Notably they identify a change point in the mean of margin of victory in 1925 related to the change in the definition of offsides (changed from 3 players to 2 players), an change points in the variability of number of goals in the early 1980s and 1992 corresponding to the change in number of points for a win (i.e. went from 2 points to 3 points) and a change in the backpass rule, respectively.  \cite{FortLee2007} looked for structural breaks in major North American sports other than baseball (i.e. basketball (NBA), hockey (NHL), and American football (NFL)).  They identified a number of change points related to competitive balance in each sport that often, though not always, correspond to league expansion, league mergers, or other major events in a sport (e.g. increased number of foreign players in the NBA in the late 1990s/early 2000s).
